NumPy---Array Programming in Python

### 2020-02-05 body.tex
Rework intro

Two Python array packages existed before NumPy.
The Numeric package began in the mid-90s and provided an array object in
Python, written in C, and linking to standard fast implementations of linear
algebra.
Around 2000, the Space Telescope Science Institute (STScI) software group wrote
a reimplementation of much of Numeric, called NumArray, to support their work
on large memory-mapped arrays and arrays of mixed data type
records \cite{STScI-slither}.
This briefly caused the two communities to diverge, until
2005, when NumPy emerged as a ``best of both worlds'' unification of Numeric
and NumArray \cite{oliphant2006guide}.

@@

Today, NumPy underpins almost every Python library that does scientific or
numerical computation including SciPy \cite{virtanen2019scipy},
Matplotlib \cite{hunter2007matplotlib}, pandas \cite{mckinney-proc-scipy-2010},
scikit-learn \cite{pedregosa2011scikit}, and
scikit-image \cite{vanderwalt2014scikit}.
It is a community-developed, open-source library, which provides a
multidimensional Python array object along with array-aware functions
that operate on it.
Because of its inherent simplicity, the NumPy array is
the \emph{de facto} exchange format for array data in Python.
The library has such widespread adoption that not only the array object but also its
\emph{Application Programming Interface} (API) has become widespread as
a language for array programming.

>>>>

Two Python array packages existed before NumPy.
The Numeric package began in the mid-1990s and provided an array object and array-aware functions
in Python, written in C, and linking to standard fast implementations of linear
algebra \cite{Numericmanual}.
To handle large astronomical images coming from the Hubble Space Telescope,
a reimplementation of Numeric, called NumArray, added
support for memory-mapped arrays and mixed data type records. 
%Around 2000, the Space Telescope Science Institute (STScI) software group wrote
%a reimplementation of Numeric, called NumArray, to support their work
%on large memory-mapped arrays and arrays of mixed data type
%records \cite{greenfield2003numarray}.
While NumArray was highly compatible with Numeric, the two packages
had enough differences that it divided the community, until
2005, when NumPy emerged as a ``best of both worlds'' unification of Numeric
and NumArray \cite{oliphant2006guide}.

@@

Today, NumPy underpins almost every Python library that does scientific or
numerical computation including SciPy \cite{virtanen2020scipy},
Matplotlib \cite{hunter2007matplotlib}, pandas \cite{mckinney-proc-scipy-2010},
scikit-learn \cite{pedregosa2011scikit}, and
scikit-image \cite{vanderwalt2014scikit}.
It is a community-developed, open-source library, which provides a
multidimensional Python array object along with array-aware functions
that operate on it.
Because of its inherent simplicity, the NumPy array is
the \emph{de facto} exchange format for array data in Python.
The library has such widespread adoption that not only the array object but also its
\emph{Application Programming Interface} (API) has become widespread as
a language for array programming.
To exploit new specialized hardware and storage, there has been
a recent proliferation of Python array packages.
Instead of leading back to some kind of Numarray and Numeric divide that
NumPy was created to bridge, NumPy has become ...
